---
title: 'CSU22013: Design and implement a dashboard to visualise and compare commit
  activity for trending repositories on GitHub.'
output:
  html_document:
    df_print: paged
---

The following chunk utilises the MIT GitHub API found at, https://github.com/alisoft/github-trending-api#readme (Full license citation to be added). 

The purpose of using this API is to create a list of the top trending repositories on GitHub, to be used later in conjunction with the official GitHub API, which will allow access to the commit data of each individual repository.

The following code firstly accesses the API, and then retrieves a list of the top trending authors and repository names, in order to manually create a list of URLs to the commit data, which the official GitHub API will then use.

A URL to a repositories commit data is as follows:
"https://api.github.com/repos/", repo_author, "/", repo_name, "/commits"

```{r}
library(httr)
library(jsonlite)
library(furrr)
library(shiny)
library(shinyjs)
library(plotly)
library(dplyr)
library(tidyr)
library(scales)
```

Function to Create base URL, to which /commits or /languages can be appended 

```{r}
API_1_create_urls <- function(date_range){
  
  # Call to MIT GitHub API to get information on current top trending repositories
  url <- "https://api.gitterapp.com/repositories?since=daily"
  response <- GET(url, query = list(since = date_range))
  repo_data <- content(response)
  
  # Retrieve list of top N authors
  repo_author <- lapply(repo_data, "[[", 1)
 
  # Retrieve list of top N repository names
  repo_name <- lapply(repo_data, "[[", 2)

# Function which manually creates URL to commit activity of each repository
  create_url <- function(repo_author, repo_name) {
  url <- paste0("https://api.github.com/repos/", repo_author, "/", repo_name)
  return(url)
  }

  # Create list of URLs to commit activity of top N repositories
  url <- mapply(create_url, repo_author, repo_name, SIMPLIFY = FALSE)
  url_list <- as.character(url)
  
  return(url_list)
}

data <- API_1_create_urls('daily')
data
```

Function which creates data frame of each repository, their languages and percentage of each language within the project

```{r}
# Personal Access Token for GitHub API
API_PAT <- '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'
headers <- c(Authorization = paste("Bearer", API_PAT))

# Function to fetch language percentages per repository
create_language_table <- function(date_range, number_of_urls) {
  language_url_list <- API_1_create_urls(date_range)
  
  # Initialise list
  language_data_list <- list()
  
  # Loop through each URL
  for (url in language_url_list[1:number_of_urls]) {
    repo_name <-
      strsplit(url, "/")[[1]][6]  # Get repository name from URL
    url <- paste0(url, "/languages")
    
    response <- GET(url, add_headers(headers))
    language_data <- content(response)
    
    # Convert language_data into a data frame
    language_df <- data.frame(
      repository = repo_name,
      language = names(language_data),
      percentage = unlist(language_data)
    )
    
    language_data_list[[repo_name]] <- language_df
  }
  
  # Combine all language dataframes into a single dataframe
  language_table <- bind_rows(language_data_list)
  
  # Pivot the data into wide format
  language_table <- language_table %>%
    pivot_wider(
      names_from = language,
      values_from = percentage,
      values_fill = 0
    )
  
  #tibble -> df
  language_table = as.data.frame(language_table)
  #colnames
  rownames(language_table) = language_table$repository
  language_table$repository <- NULL
  
  # Calculate the row sums
  row_sums <- rowSums(languageDataFrame)

  # Calculate the percentage of each column by row to 2 decimal places
  round(languageDataFrame / row_sums, 2)
  
  return(language_table)
}

languageDataFrame <- create_language_table('daily', 10)

```

```{r}
rownames(language_data) = language_data$repository
```


Function which creates a data frame of every commit in the master branch of a repo. Each commit in data frame includes: Repository name, developer name, developer contact email, datetime commit published at

```{r}
# Personal Access Token for GitHub API
API_PAT <- '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'
headers <- c(Authorization = paste("Bearer", API_PAT))

#date_range = 'daily'
#number_of_urls = 1

url = commit_activity_url_list[[1]]


create_commits_table <- function(date_range, number_of_urls) {
  commit_activity_url_list <- API_1_create_urls(date_range)
  
  # Initialize an empty data frame to store the commit information
  cols = c('repo','dev_name','email','datetime')
  data_table = data.frame(matrix(nrow = 0, ncol = length(cols))) 
  colnames(data_table) = cols
  
  # Loop through each commit activity URL
  for (url in commit_activity_url_list[1:number_of_urls]) {
    
    url <- paste0(url, "/commits")
    
    page <- 1

    while (TRUE) {
      page_url <- paste0(url, "?per_page=100&page=", page)
      response <- GET(page_url, add_headers(headers))
      commit_data <- content(response)
      
      if (length(commit_data) == 0) {
        break  # No more commits on this page
      }
      
      df <- data.frame(
        repo = c(strsplit(unlist(lapply(commit_data, function (x) x$commit$tree$url)), "/")[[1]][6]),
        dev_name = c(unlist(lapply(commit_data, function (x) x$commit$author$name))),
        email = c(unlist(lapply(commit_data, function (x) x$commit$author$email))),
        datetime = c(unlist(lapply(commit_data, function (x) x$commit$author$date)))
      )
      
      data_table <- rbind(data_table, df) 
      page <- page + 1
    }
  
  }
  
  return(data_table)
}

# Call the function with desired parameters
commitDataFrame <- create_commits_table('daily', 3)
commitDataFrame
```

The following chunk concerns the official GitHub API, found at, https://docs.github.com/en/rest?apiVersion=2022-11-28.

Unfortunately there is no header within the API for any kind of 'total commit count', however it is possible to search page-by-page through the commit history. The maximum commits which can be displayed per page is 100. In order to retrieve the total number of commits per repository, I created a function which loops through each page of the commit history, counting up to 100 commits at a time, until there are no more commits left to count. The function does this for the top 10 trending repositories, the list for which was created in the previous chunk. 

Initially, I was calculating the total commits for the top 25 repositories, however this was taking approximately 3 minutes to complete, so I decided to only analyse the top 10 repositories.

The process in my testing takes approximately 30 seconds, however depending on the total number of commits within the top 10 repositories, this could take a longer, or shorter length of time.



```{r}
fetch_total_commits <- function(commit_activity_url) {
  # Set page number to 1, and total commit count to 0
  page <- 1
  total_commits <- 0
  
  # Set max commits shown per page to 100, and loop through each page, counting total number of commits, until there are none left to count
  while (TRUE) {
    commits_url <- paste0(commit_activity_url, "?per_page=100&page=", page)
    response <- GET(commits_url, add_headers(headers))
    commit_data <- content(response)
    
    # Calculate number of commits on page
    commits_on_page <- length(commit_data)
    
    # If there are no commits in repository, exit loop and return total commits as 0
    if (commits_on_page == 0) {
      break
    }
    
    # Add total commits from page to total commit count, and go to next page
    total_commits <- total_commits + commits_on_page
    page <- page + 1
  }
  
  # Return value of total repository commits
  return(total_commits)
}
```

The following chunk utilises Shiny, an open source package for R, to build the app which creates and displays the commit data.

At the moment the code creates and displays a sample bar plot, comparing the total number of commits of the top 10 trending repositories. 

total_day_commits_list <- future_map_int(commit_activity_url_list, fetch_total_day_commits)

total_night_commits_list <- future_map_int(commit_activity_url_list, fetch_total_night_commits)

```{r}
# Define UI for app which draws barplot
ui <- fluidPage(
  # App title
  titlePanel("Top 5 Trending GitHub Repositories"),
  
  # Drop-down input for time period selection
  selectInput(
    inputId = "time_period",
    label = "Select Time Period:",
    choices = c("Total", "Last 24 Hours", "Last Week", "Last Month", "Last 6 Months", "Last Year"),
    selected = "Total"
  ),
  
  # Output: barplot
  shinycssloaders::withSpinner(plotlyOutput(outputId = "bar_plot"))
)

# Define server logic required to draw barplot
server <- function(input, output) {
  # Render the bar chart
  output$bar_plot <- renderPlotly({
    # Set up parallel processing
    plan(multisession, workers = availableCores())
    
    # Select the appropriate total_commits list based on user selection
    selected_list <- switch(
      input$time_period,
      "Total" = total_commits_list <-
        future_map_int(commit_activity_url_list, fetch_total_commits),
      "Last 24 Hours" = total_commits_since_24_hours_list <-
        future_map_int(commit_activity_url_list, fetch_total_commits_24hours),
      "Last Week" = total_commits_since_last_7_days_list <-
        future_map_int(
          commit_activity_url_list,
          fetch_total_commits_last_7_days
        ),
      "Last Month" = total_commits_since_last_month_list <-
        future_map_int(
          commit_activity_url_list,
          fetch_total_commits_last_month
        ),
      "Last 6 Months" = total_commits_since_last_6_months_list <-
        future_map_int(
          commit_activity_url_list,
          fetch_total_commits_last_6_months
        ), 
      "Last Year" = total_commits_since_last_year_list <-
        future_map_int(
          commit_activity_url_list,
          fetch_total_commits_last_year
        )
    )
    
    # Create the bar chart using barplot
    plot_ly(
      x = repo_name,
      y = selected_list,
      type = "bar",
      marker = list(color = "#007bc2"),
      hoverinfo = paste(selected_list)
    ) %>%
      layout(
        xaxis = list(title = "Repository Names in Descending Order from Most Trending to Least Trending"),
        yaxis = list(title = "Total Commits"),
        title = "Commits for the Top Trending Repositories on GitHub"
      )
    
  })
  
}

# Create Shiny app
shinyApp(ui = ui, server = server)
```
