---
title: 'CSU22013: Design and implement a dashboard to visualise and compare commit
  activity for trending repositories on GitHub.'
output:
  html_document:
    df_print: paged
---

The following chunk utilises the MIT GitHub API found at, https://github.com/alisoft/github-trending-api#readme (Full license citation to be added). 

The purpose of using this API is to create a list of the top trending repositories on GitHub, to be used later in conjunction with the official GitHub API, which will allow access to the commit data of each individual repository.

The following code firstly accesses the API, and then retrieves a list of the top trending authors and repository names, in order to manually create a list of URLs to the commit data, which the official GitHub API will then use.

A URL to a repositories commit data is as follows:
"https://api.github.com/repos/", repo_author, "/", repo_name, "/commits"

```{r}
# Libraries
library(httr)
library(jsonlite)
library(furrr)
library(shiny)
library(shinyjs)
library(plotly)
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(DT)
library(shinycssloaders)

# Personal Access Token for GitHub API
API_PAT <<- c(Authorization = paste("Bearer", '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'))
PAGE_LIMITS <<- 5

# default values for selections in left menu
# daterange = 'daily'
# language = 'all'

```

Function to Create base URL, to which /commits or /languages can be appended. Daterange fetches either top repositories from last day, week, or month.

```{r}
# Function: Get list of base URLs for trending repositories
API_1_create_urls <- function(date_range){
  #print('###################### API_1 ######################')
  # Call to MIT GitHub API to get information on current top trending repositories
  url <- "https://api.gitterapp.com/repositories?since=daily"
  response <- GET(url, query = list(since = date_range))
  repo_data <- content(response)
  
  # Retrieve list of trending authors
  repo_author <- lapply(repo_data, "[[", 1)
 
  # Retrieve list of trending repository names
  repo_name <- lapply(repo_data, "[[", 2)

# Function: Manually create base URL to repository
  create_url <- function(repo_author, repo_name) {
  url <- paste0("https://api.github.com/repos/", repo_author, "/", repo_name)
  return(url)
  }

  # Create list of URLs to top repositories
  url <- mapply(create_url, repo_author, repo_name, SIMPLIFY = FALSE)
  url_list <- as.character(url)
  
  return(url_list)
}
```

Function which creates data frame of each repository, their languages and percentage of each language within the project

```{r}
# languageDataFrame
# Function to fetch language percentages per repository
create_language_dataframe <- function() {
  #print('###################### API_2 ######################')
  # Initialise list
  language_data_list <- list()
  
  # Loop through each URL
  for (url in GIT_URLS) {

    repo_name <- strsplit(url, "/")[[1]][6]  # Get repository name from URL
    url <- paste0(url, "/languages")         # Append /languages to base URL
      
    response <- GET(url, add_headers(API_PAT))
    language_data <- content(response)

    if(length(language_data) == 0){
      GIT_URLS = GIT_URLS[-which(GIT_URLS == url)]
      next
    }
    
    # Convert language_data into a data frame
    language_df <- data.frame(
      repository = repo_name,
      language = names(language_data),
      percentage = unlist(language_data)
    )
    
    language_data_list[[repo_name]] <- language_df
  }
  
  # Combine all language dataframes into a single dataframe
  language_table <- bind_rows(language_data_list)
  
  # Pivot the data into wide format
  language_table <- language_table %>%
    pivot_wider(
      names_from = language,
      values_from = percentage,
      values_fill = 0
    )
  
  # Tibble -> df
  language_table = as.data.frame(language_table)
  # Colnames
  rownames(language_table) = language_table$repository
  language_table$repository <- NULL
  
  # Calculate the row sums
  row_sums <- rowSums(language_table)

  # Calculate the percentage of each column by row to 2 decimal places
  language_table <- round(language_table / row_sums, 2)
  
  #Remove languages that contribute less than 1% of the code
  language_table <- language_table[,colSums(language_table != 0) > 0]
  
  return(language_table)
}
```

Function which filters languageDataFrame by programming language, returning the top 5 (or less if there are less than 5) repositories which contain the user specified language (e.g. Python)

```{r}
# filteredLanguageDataFrame
# Function: Create data frame to display the top 5 trending repositories, filtering by programming language. "All" returns top 5 generally
create_filtered_top_n_language_dataframe <- function(language, language_df, n = 5) {
  # print('create_filtered_top_n_language_dataframe')
  if (language == "All") {
      # Select the top n repositories based on trendiness
      return(head(language_df, n))
  }
  
  # Get the index of the chosen language in the data frame
  language_index <- which(colnames(language_df) == language)
    
  # Filter out repositories with zero% of chosen language
  filtered_repositories <- language_df[language_df[, language_index] > 0, ]
    
  # Select repositories, considering the case of less than n repositories
  num_repositories <- min(nrow(filtered_repositories), n)
  top_repositories <- filtered_repositories[1:num_repositories, ]

  return(top_repositories)
}
```

Function which creates a data frame of every commit in the master branch of a repo. Each commit in data frame includes: Repository name, developer name, developer contact email, datetime commit published at

```{r}
# commitDataFrame
# Function: Get the entire commit history of n number of repositories from top trending
create_commits_dataframe <- function() {
  # print('create_commits_dataframe')
  # Initialize an empty data frame to store the commit information
  cols = c('repo','dev_name','email','datetime')
  data_table = data.frame(matrix(nrow = 0, ncol = length(cols))) 
  colnames(data_table) = cols
  
  # Loop through each commit activity URL
  for (url in GIT_URLS_FILTERED) {
    
    # Append /commits to base URL
    url <- paste0(url, "/commits")
    
    page <- 1

    while (page <= PAGE_LIMITS) {
      page_url <- paste0(url, "?per_page=100&page=", page)
      response <- GET(page_url, add_headers(API_PAT))
      commit_data <- content(response)
      
      # Break if no more commits
      if (length(commit_data) == 0) {
        break  
      }
      
      df <- data.frame(
        repo = c(strsplit(unlist(lapply(commit_data, function (x) x$commit$tree$url)), "/")[[1]][6]),
        dev_name = c(unlist(lapply(commit_data, function (x) x$commit$author$name))),
        email = c(unlist(lapply(commit_data, function (x) x$commit$author$email))),
        datetime = c(unlist(lapply(commit_data, function (x) x$commit$author$date)))
      )
      
      # Add commit data to main data frame
      data_table <- rbind(data_table, df) 
      page <- page + 1
    }
  
  }
  
  return(data_table)
}

```

```{r}
# Function: Retrieve a list of all of the languages used in top repositories
get_list_of_languages <- function(language_df){
  #print('get_list_of_languages')
  return(colnames(language_df))
  
}
```


```{r}
# Function: Get the URLs from list of repositories
get_urls_from_list_of_repo_names  = function(language_df){
   #print('get_urls_from_list_of_repo_names')
    filtered_urls <- list()
    for(url in GIT_URLS){
      name <- strsplit(url, "/")[[1]][6]
      if(name %in% rownames(language_df)){
        filtered_urls[name] <- url
      }
    }  
    
  return(filtered_urls)
}
```


```{r}
# Function: Count of all commits grouped by repo and then developer. Email for each developer included
count_commits_by_DevRepo <- function(commits_df){
    # print('count_commits_by_DevRepo')
    aggregatedDataFrame = aggregate(email ~ dev_name + repo, data = commits_df, FUN = length)
    aggregatedDataFrame = setNames(aggregatedDataFrame, c('dev_name', 'repo', 'count'))
    aggregatedDataFrame = merge(x = aggregatedDataFrame, y = commits_df[,c('dev_name','email')], by = 'dev_name')
    
    aggregatedDataFrame <- aggregatedDataFrame[!duplicated(aggregatedDataFrame[c("dev_name", "repo", "count", "email")]), ]
    
    aggregatedDataFrame <- aggregatedDataFrame[order(aggregatedDataFrame$repo, -aggregatedDataFrame$count),]
    
    return(aggregatedDataFrame)
}

```

```{r}
# Function: Get the top n developers, per top n repository
top_n_commiters_per_repo <- function(commits_count, n = 5) {
  # print('top_n_commiters_per_repo')
  # Get unique repos 
  unique_repos <- unique(commits_count$repo)
  
  # Create new data frame to store top n data
  result_df <- data.frame(dev = character(0), repo = character(0), counts = numeric(0))
  
  # Loop through each repository
  for (repo in unique_repos) {
    repo_data <- commits_count[commits_count$repo == repo, ]  # Get rows for each repo
    top_devs <- repo_data[order(-repo_data$count), ][1:min(dim(repo_data)[1],n), ] # Order by count and get top n devs
    result_df <- rbind(result_df, top_devs) # Add results to new data frame
  }
  
  row.names(result_df) <- 1:nrow(result_df)
  

  
  return(result_df)
}
```

```{r}
# Function: Create a percentage stacked bar plot for total languages used within top n repositories
graph_languages <- function(df){
  
  # print('graph_languages')
  # Find columns where all values are 0
  cols_to_remove <- colnames(df)[apply(df, 2, function(col) all(col == 0))]

  # Remove zero value columns
  df_filtered <- df[, !(colnames(df) %in% cols_to_remove)]

  language_matrix <- t(as.matrix(df_filtered))

  # Get remaining repository names
  repo_names <- rownames(df_filtered)

  stacked_barplot <- plot_ly()

  # Add a trace for each programming language
  for (i in 1:nrow(language_matrix)) {
    stacked_barplot <- stacked_barplot %>%
      add_trace(
        x = repo_names,
        y = language_matrix[i, ],
        type = "bar",
        name = rownames(language_matrix)[i]
      )
  }

  # Generate bar plot
  stacked_barplot <- stacked_barplot %>%
    layout(
      title = "Percentage Breakdown of Languages by Top 5 Repository",
      xaxis = list(title = "Repository"),
      yaxis = list(title = "Percentage"),
      barmode = "stack"
    )
  
  return(stacked_barplot)
}



```

```{r}
# Function: Create a grouped barplot of top n developers for top n repos, commit counts

graph_commits <- function(df) {
   # print('graph_commits')
plot <- plot_ly(df, 
                x = ~repo, 
                y = ~count, 
                text = ~paste("Developer: ", dev_name, "<br>No. Commits: ", count), 
                hoverinfo = 'text', 
                type = 'bar', 
                split = ~dev_name) %>%
  layout(
    title = 'Number of Commits per Top 5 Developer by Repository',
    xaxis = list(title = 'Repository'),
    yaxis = list(title = 'Number of Commits'),
    barmode = 'group'
  )
return(plot)
}

```


```{r}
# Function: Create an information table to display referenced developers information - name, repository, email address

create_dev_info_table <- function(df){
  
  return(datatable(df[, -which(names(df) == "count")],options = list(scrollY = "300px")))

}
```


PLOTTERS 

```{r}
# Function: Retrieves data for programming languages percentage stacked bar chart
stackedLanguageBarplot <- function(language_df_filtered) {
  
  return(graph_languages(language_df_filtered)) 

}
```

```{r}
# Function: Retrieves data for developer information table
dev_info_table <- function(commits_count) {

  return(create_dev_info_table(commits_count))
  
}
```


```{r}
# Function: Retrieves data for commit count bar plot graph
groupedCommitsBarplot <- function(commits_count) {
  
  return(graph_commits(commits_count))
  
}
```


EVENT HANDLERS

```{r}
date_range_event <- function(date_range){
    GIT_URLS <<- API_1_create_urls(date_range)
    # Create data frame of repositories, programming languages, and language percentages
    LANGUAGE_DF <<- create_language_dataframe()
}

language_event <- function(input, output, language){
      
      language_df_filtered = create_filtered_top_n_language_dataframe(language, LANGUAGE_DF)
      
      GIT_URLS_FILTERED <<- get_urls_from_list_of_repo_names(language_df_filtered)
  
      commits_df = create_commits_dataframe()
      
      # Count number of commits for each developer
      commits_count = count_commits_by_DevRepo(commits_df)
          
      # Calculate top 5 developers for each repo, based on number of commits
      commits_count = top_n_commiters_per_repo(commits_count)
      
      # Display stacked language bar plot
      output$stackedBarplot <- renderPlotly({
        stackedLanguageBarplot(language_df_filtered)
      })
      
      # Display developer info table 
      output$table <- renderDT({
        dev_info_table(commits_count)
      })
    
      # Display grouped commits bar plot
      output$groupedBarplot <- renderPlotly({
        groupedCommitsBarplot(commits_count)
      })
      
      return(output)
}
```


```{r}
# Function: Populates drop-down list
getLanguages <- function() {
      
      dropdown_list <- sort(get_list_of_languages(LANGUAGE_DF))
      dropdown_list <- c("All", dropdown_list)
      
      return(dropdown_list)
}
```


Init LANGUAGE_DF for the languages dropdown
```{r}
date_range_event('daily')
```

SERVER / UI FRAMEWORK

```{r}
ui <- shinyUI(fluidPage(

  titlePanel("Trending Repositories on GitHub"),
    
  sidebarLayout(
    # Date range drop-down menu
    sidebarPanel(width = 2, selectInput(
      "date_range_dropdown",
      "GitHub Date Range:",
      choices = c("Daily", "Weekly", "Monthly")
    ),
    selectInput(
      # Language drop-down menu
      "language_dropdown",
      "Programming Language:",
      choices = getLanguages()
    )),
    mainPanel(
      withSpinner(plotlyOutput("groupedBarplot")),
      withSpinner(plotlyOutput("stackedBarplot")),
      withSpinner(dataTableOutput("table"))
    )
  )
))


# Define server function
server <- shinyServer(function(input, output) {
  
  observeEvent(input$date_range_dropdown, {
    date_range_event(input$date_range_dropdown)
    output <- language_event(input, output, "All")
  })

  observeEvent(input$language_dropdown,{
    output <- language_event(input, output, input$language_dropdown)
  })
  
 })

# Run the Shiny app
shinyApp(ui, server)
```

